Calendar-based graphics for visualising people's daily schedules

### 2017-07-06 main.Rmd
roughly done with grid construction

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Di Cook
    affiliation: Monash Univeristy
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash Univeristy
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualising people's daily schedules"
  plain: "Calendar-based graphics for visualising people's daily schedules"
abstract: >
  It is often challenging to visualise people's daily schedules as the data are of multiple seasons and of long length. One such example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations. Owing to human-related activities, the primary depiction of the data features multiple temopral scales such as time of day, day of week, day of year. We introduce the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail. We provide some examples to demonstrate its usage by applying to the pedetrian data.
keywords:
  # at least one keyword must be supplied
  formatted: [calendar-based visualisation, temporal-context data]
  plain:     [calendar-based visualisation, temporal-context data]
preamble: >
  \usepackage{amsmath}
output: 
  rticles::jss_article:
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

* $k = 1, \dots , 7$ is the day of the week that is the first day of the month
* $d = 28, 29, 30$ or $31$ representing the number of days in any month
* $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week. 
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells

@@

Each cell forms a canvas on which to draw the data. Consider the canvas to have limits $[0, 1]$ horizontally and vertically. For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically. Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed. Let $h$ be the scaled hour, and $c$ the scaled count.

@@

The algorithm can be extended relatively easily to layout from one single month to multiple years depending on the time span to be visualised. The month-by-month arrangement can be determined by the user's choice through $M$ and $N$. If one would like to compare and contrast January, February and so on across multiple years, $M = 12$ and $N$ depending on the number of years could be setup.

Monthly calendar format provides an effective layout to emphasise differences between weekdays and weekends as well as day of year (for example, public holiday) simultaneously, however, either one of which is sufficient in certain visualisation applications. If there is apparent depiction of weekday effects but none of yearly effects, weekly calendar format (weeks of a year) could be considered employing; otherwise days of a month could be laid out for standing-out yearly behaviours.

(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE). Weekly and daily calendar formats are the simplifications of monthly calendar.

We only illustrate that calendar grids are laid out on the horizontal direction. The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above. It is useful for users in some countries where they get used to calendars of vertical organisation and other possible comparisons like Mondays across the columns rather than rows.

>>>>

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Di Cook
    affiliation: Monash Univeristy
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash Univeristy
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualising people's daily schedules"
  plain: "Calendar-based graphics for visualising people's daily schedules"
abstract: >
  It is challenging to visualise people's daily schedules as such data are often of multiple seasons and of long time span. One example is Melbourne pedestrian sensor data which are collected at hourly interval and at a number of locations. Owing to human-related activities, the primary depiction of the data features multiple temopral scales such as time of day, day of week, day of year. We develop the \code{frame\_calendar} function that provides a method for rearranging such kind of data in a calendar-based format and accomplishing visualisation with the grammar of graphics, and describe its construction in detail. We provide some examples to demonstrate its usage by applying to the pedetrian data.
keywords:
  # at least one keyword must be supplied
  formatted: [calendar-based visualisation, temporal-context data]
  plain:     [calendar-based visualisation, temporal-context data]
preamble: >
  \usepackage{amsmath}
output: 
  rticles::jss_article:
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

* $k = 1, \dots , 7$ is the day of the week that is the first day of the month.
* $d = 28, 29, 30$ or $31$ representing the number of days in any month.
* $(i, j)$ is the grid position where $1 \le i \le 5$ is week within the month, $1 \le j \le 7$, is day of the week. 
* $g = k, \dots,(k+d)$ indexes the day in the month, inside the 35 possible cells.

@@

Each cell forms a canvas on which to draw the data. Consider the canvas to have limits $[0, 1]$ horizontally and vertically. For the pedestrian sensor data, within each cell hour is plotted horizontally and count is plotted vertically. Each variable is scaled to have values between $[0,1]$, using the minimum and maximum of all the data values to be displayed assuming of fixed scales. Let $h$ be the scaled hour, and $c$ the scaled count.

@@

The algorithm can be relatively easily extended to layout from a single month to a few years, based on the period of time to be visualised. $(M, N)$ can be determined by the user using the arguments of \code{nrow} and \code{ncol}. If the one would like to visualise data spanning over three years, for example, $M = 12$ and $N = 3$ seem to be an appropriate choice when assessing the differences of a given month across the years.

The algorithm can be simplified to accommodate the other two types of calendar formats. One is comprised of days of a week in columns and weeks of a year in rows, and the other is days of a month in columns and months of a year in rows, which we refer to as "weekly" and "daily" calendar respectively. Due to the arrangements, the weekly calendar puts more emphases on days of a week over days of a year, whereas the daily one serves as the opposite. The monthly format can be considered as the advanced twist of both weekly and daily. Temporal patterns lead to which format to be employed. The weekly calendar can be a nice attempt if the most variations can be characterised by days of a week. On the other hand, the absence of weekly effect but the presence of yearly effect can direct to the daily calendar. When both are present, the monthly calendar appears to be a better choice.

We only illustrate that grids are laid out over the horizontal direction in the paper. The vertical direction can be enabled by swapping $i$ and $j$ in the algorithm stated above. It is particularly useful for those users who get accustomed to calendars of vertical organisation in some countries.

### 2017-07-06 main.tex
roughly done with grid construction

\Abstract{
It is often challenging to visualise people's daily schedules as the
data are of multiple seasons and of long length. One such example is
Melbourne pedestrian sensor data which are collected at hourly interval
and at a number of locations. Owing to human-related activities, the
primary depiction of the data features multiple temopral scales such as
time of day, day of week, day of year. We introduce the
\code{frame\_calendar} function that provides a method for rearranging
such kind of data in a calendar-based format and accomplishing
visualisation with the grammar of graphics, and describe its
construction in detail. We provide some examples to demonstrate its
usage by applying to the pedetrian data.
}

@@

\begin{itemize}
\tightlist
\item
  \(k = 1, \dots , 7\) is the day of the week that is the first day of
  the month
\item
  \(d = 28, 29, 30\) or \(31\) representing the number of days in any
  month
\item
  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within
  the month, \(1 \le j \le 7\), is day of the week.
\item
  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35
  possible cells
\end{itemize}

@@

Each cell forms a canvas on which to draw the data. Consider the canvas
to have limits \([0, 1]\) horizontally and vertically. For the
pedestrian sensor data, within each cell hour is plotted horizontally
and count is plotted vertically. Each variable is scaled to have values
between \([0,1]\), using the minimum and maximum of all the data values
to be displayed. Let \(h\) be the scaled hour, and \(c\) the scaled
count.

@@

The algorithm can be extended relatively easily to layout from one
single month to multiple years depending on the time span to be
visualised. The month-by-month arrangement can be determined by the
user's choice through \(M\) and \(N\). If one would like to compare and
contrast January, February and so on across multiple years, \(M = 12\)
and \(N\) depending on the number of years could be setup.

Monthly calendar format provides an effective layout to emphasise
differences between weekdays and weekends as well as day of year (for
example, public holiday) simultaneously, however, either one of which is
sufficient in certain visualisation applications. If there is apparent
depiction of weekday effects but none of yearly effects, weekly calendar
format (weeks of a year) could be considered employing; otherwise days
of a month could be laid out for standing-out yearly behaviours.

(MAKE TWO DIAGRAMS ON WEEKLY AND DAILY CALENDAR FORMATS IN KEYNOTE).
Weekly and daily calendar formats are the simplifications of monthly
calendar.

We only illustrate that calendar grids are laid out on the horizontal
direction. The vertical direction can be enabled by swapping \(i\) and
\(j\) in the algorithm stated above. It is useful for users in some
countries where they get used to calendars of vertical organisation and
other possible comparisons like Mondays across the columns rather than
rows.

>>>>

\Abstract{
It is challenging to visualise people's daily schedules as such data are
often of multiple seasons and of long time span. One example is
Melbourne pedestrian sensor data which are collected at hourly interval
and at a number of locations. Owing to human-related activities, the
primary depiction of the data features multiple temopral scales such as
time of day, day of week, day of year. We develop the
\code{frame\_calendar} function that provides a method for rearranging
such kind of data in a calendar-based format and accomplishing
visualisation with the grammar of graphics, and describe its
construction in detail. We provide some examples to demonstrate its
usage by applying to the pedetrian data.
}

@@

\begin{itemize}
\tightlist
\item
  \(k = 1, \dots , 7\) is the day of the week that is the first day of
  the month.
\item
  \(d = 28, 29, 30\) or \(31\) representing the number of days in any
  month.
\item
  \((i, j)\) is the grid position where \(1 \le i \le 5\) is week within
  the month, \(1 \le j \le 7\), is day of the week.
\item
  \(g = k, \dots,(k+d)\) indexes the day in the month, inside the 35
  possible cells.
\end{itemize}

@@

Each cell forms a canvas on which to draw the data. Consider the canvas
to have limits \([0, 1]\) horizontally and vertically. For the
pedestrian sensor data, within each cell hour is plotted horizontally
and count is plotted vertically. Each variable is scaled to have values
between \([0,1]\), using the minimum and maximum of all the data values
to be displayed assuming of fixed scales. Let \(h\) be the scaled hour,
and \(c\) the scaled count.

@@

The algorithm can be relatively easily extended to layout from a single
month to a few years, based on the period of time to be visualised.
\((M, N)\) can be determined by the user using the arguments of
\code{nrow} and \code{ncol}. If the one would like to visualise data
spanning over three years, for example, \(M = 12\) and \(N = 3\) seem to
be an appropriate choice when assessing the differences of a given month
across the years.

The algorithm can be simplified to accommodate the other two types of
calendar formats. One is comprised of days of a week in columns and
weeks of a year in rows, and the other is days of a month in columns and
months of a year in rows, which we refer to as ``weekly'' and ``daily''
calendar respectively. Due to the arrangements, the weekly calendar puts
more emphases on days of a week over days of a year, whereas the daily
one serves as the opposite. The monthly format can be considered as the
advanced twist of both weekly and daily. Temporal patterns lead to which
format to be employed. The weekly calendar can be a nice attempt if the
most variations can be characterised by days of a week. On the other
hand, the absence of weekly effect but the presence of yearly effect can
direct to the daily calendar. When both are present, the monthly
calendar appears to be a better choice.

We only illustrate that grids are laid out over the horizontal direction
in the paper. The vertical direction can be enabled by swapping \(i\)
and \(j\) in the algorithm stated above. It is particularly useful for
those users who get accustomed to calendars of vertical organisation in
some countries.
